Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2019-07-13 unimodular-covers.tex
checking comments

\begin{corollary}
\label{coro:coverpara}
\paco{added this intermediate result}
\giulia{We are using lemmas 2 and 3 which are for empty tetrahedra; either we add a step in this corollary saying that we can reduce to empty tetrahedra, or we keep this for empty and triangulate into empty tetrah in the proof of the theorem (which makes more sense to me)}
Let $T$ be a empty lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
\end{corollary}

@@

\section{ Cayley sums}
\label{sec:cayley}

Throughout this section \giulia{I think this may be misleading, in the sense that later we specify to $d=2$ and we should maybe highlight this} let $P$ and $Q$ be two lattice polytopes in $\R^d$. We do not require them to be full-dimensional, but we assume their Minkowski sum is. Remember that the Minkowski sum $P+Q$ of $P$ and $Q$ is defined as:
\[
P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.
\]
The \emph{Cayley sum} of $P$ and $Q$ is defined as
\[
\cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \in \R^{d+1}.
\]
The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ being $(d+1)$-dimensional.

@@

@@

\begin{proof}[Proof of \Cref{thm:cayley}]
Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$. Observe that any simplex in $T$ has vertices in $Px\{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$. Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$). \giulia{is it usually denoted by dimension or cardinality? that is, (2,0) or (3,1)?} A triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained by simply unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$). This leaves simplices of type $(2,2)$. In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.

@@

\begin{claim}\giulia{I think we need to treat separately the case where we deal with edges (these are easier, in any case, but I couldn't find a unification)}
The following inequalities hold, 
\begin{align*}
\width_{f_q}(\ell) , 
\width_{f_q}(r) ,
\width_{f_p}(t) ,
\width_{f_p}(b) \geq w.
\end{align*}
Each inequality is strict, unless the segment in question coincides with $p$ or $q$.
\end{claim}

@@

Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
\end{proof}
\giulia{add a picture!}

>>>>

\begin{corollary}
\label{coro:coverpara}
Let $T$ be a empty lattice tetrahedron contained in a lattice parallelepiped $P$. Then, $T$ can be covered by unimodular tetrahedra contained in $P$.
\end{corollary}

@@

\section{ Cayley sums} \giulia{Change notations: remove nef and make sure everything works}
\label{sec:cayley}
Let $P$ and $Q$ be two lattice polytopes in $\R^d$. We do not require them to be full-dimensional, but we assume their Minkowski sum is. Remember that the Minkowski sum $P+Q$ of $P$ and $Q$ is defined as:
\[
P + Q := \{ p+q \in \R^d: p\in P, q \in Q\}.
\]
The \emph{Cayley sum} of $P$ and $Q$ is defined as
\[
\cayley (P,Q) = \conv( P\times\{0\} \cup Q\times \{1\}) \in \R^{d+1}.
\]
The assumption that $P+Q$ is $d$-dimensional is equivalent to $\cayley (P,Q)$ being $(d+1)$-dimensional.

@@

\medskip
\giulia{move first paragraph of proof of theorem here, and explain simplices, prodsimplices...}

We now turn our attention to dimension $d=2$.  The unimodular prod-simplices are unimodular parallelograms $p+q$ with $p$ a segment of $P$ and $q$ a segment of $Q$, and unimodular triangles $T+u$, where $T \subset P$ and $u \in Q$ or viceversa.

Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$. Observe that any simplex in $T$ has vertices in $P \times \{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$. Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$).

\giulia{fix this}

@@

\begin{proof}[Proof of \Cref{thm:cayley}]

A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$). This leaves simplices of type $(2,2)$. In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.

@@

\begin{claim}
The following inequalities hold, 
\begin{align*}
\width_{f_q}(\ell) , 
\width_{f_q}(r) ,
\width_{f_p}(t) ,
\width_{f_p}(b) \geq w.
\end{align*}
Each inequality is strict, unless the segment in question coincides with $p$ or $q$.
\end{claim}

@@

Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
\end{proof}
\giulia{add a picture?}

\Giulia{add references to Haase&Hofmann and Tsuchiya}
